\documentclass[ENG,PhD]{lti-tech-report}
%\documentclass[ENG,MSc,DRAFT]{cinvestav}
\usepackage{color}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{latexsym, amssymb}
\usepackage{xcolor}
\usepackage{pstricks}
\usepackage{epsfig}
\usepackage{ucs}
\usepackage{lettrine}
\usepackage[utf8x]{inputenc}
\usepackage{colortbl}
\usepackage{url}
%\usepackage{pseudocode}
\usepackage{longtable}
\RequirePackage{hyperref}
\RequirePackage{algorithm}
\usepackage{algorithmic}
\usepackage{latexsym, amssymb}
\usepackage{subfigure}
\usepackage{fancybox}
\usepackage{rotating}
\usepackage{pdflscape} 
\usepackage{multirow}
\usepackage{textcomp}
\floatname{algorithm}{Procedure}
\newcommand{\V}[1]{\mathbf{#1}}
%\graphicspath{{figs/}}
%\usepackage[english]{babel}\selectlanguage{english}
%\usepackage[spanish]{babel}\selectlanguage{spanish}
\title{Security of Data in Domotic Applications with Wireless Sensor Networks}
\shorttitle {Security of Data in Smart House Applications with WSN}
\author {Brisbane Ovilla Martínez}
\adscripcion{CINVESTAV UNIDAD TAMAULIPAS. LABORATORIO DE TECNOLOGÍAS DE INFORMACIÓN. Parque Científico y Tecnológico TECNOTAM -- Km. 5.5 carretera Cd. Victoria-Soto La Marina. C.P. 87130 Cd. Victoria, Tamps. }
\city {Cd. Victoria, Tamaulipas, México.}
%\date{December, 2010}
\techreportnumber{1}
\publishedmonth{April} %mes de la publicación
\publishedday{30} %día de la publicación
\publishedyear{2012} %año de la publicación
\keywords{Key, PUF, domotic, }
\correspondingauthor{Brisbane Ovilla Martínez $<$brisbane@tamps.cinvestav.mx$>$}
%\grants{Si recibieron financiamiento, agréglenlo, por ejemplo: This research was partially funded by project number 51623 from ``Fondo Mixto Conacyt-Gobierno del Estado de Tamaulipas''}
\dateofsubmission{Mayo 2, 2012}
\tobecited{Brisbane1}
\placeanddateofpublication{Ciudad Victoria, Tamaulipas, MEXICO.}
\abstract{
Wireless sensor networks (WSN) are a technology that in recent years has had a lot of interest by researchers. Can implement an unlimited number of applications, including application in home automation. The use of sensor networks in a house has strong implications for security as the information passing through the network is of a personal nature, so it is important to add security mechanisms to protect information. One of the security issues addressed by researchers is the establishment of keys in WSN, there are different proposal for the type of communication that must occur between the network elements. In this paper we study in detail the key establishment schemes that are applied to home automation and based on them, we show some proposals of the same mechanisms but with the integration of the use of PUF.

}
\begin{document}
\makeintropages
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Introduction
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

The wireless sensor network (WSN) consist of many spatially distributed devices (called nodes) using sensors to detect and monitor various conditions as shown in Figure \ref{fig:sistema}. Nodes collect and submit data either through other nodes or to a base station.  The base station may be a node with more computing power and/or direct connection to any infrastructure where the user can access data \cite{Al-Karaki:2004uq}.



\newpage
%\bibliographystyle{abbrv}
%\bibliographystyle{unsrt}
\bibliographystyle{apalike}
\bibliography{../../../Bibliografia/bibliografia}
\end {document}
